The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-17 paper/ref_approach.tex
fixed section 3: references and other things

In this section, we demonstrate the benefits of reference model approaches using the normal means problem as a benchmark. The normal means problem consists of estimating the (usually sparse) vector of means of a vector of normally distributed observations. Let us denote the dimensionality of the vector of means by $p$ and let $\{z_{j}\}_{j=1}^{p}$ be the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}) \quad j=1,..,p
\end{equation}
Note that it is equivalent to a linear regression where the design matrix is the identity matrix with the number of features being equal to the number of observations. This formulation can be found in practice at various places, 
for example, in the analysis of microarray data, where a large set of genes are tested in two groups of patients labeled as positive or negative to some disease (cite). The objective of the analysis is to select the subset of genes statistically relevant to the disease. One common way to proceed is to compute the two-samples $t$-statistic for every gene separately.
After normalizing these statistics, they then become the data, $Z$, in the normal means problem \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}.

@@

We begin by explaining some Bayesian estimation methods to solving the normal means problem. They all have in common that a Bayesian regression model is fitted using some sparsifying prior on the regression coefficients, for example, as discussed in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. If these priors have continuous support, as is the case for the Dirichlet-Laplace (DL; cite) or horseshoe-like priors (cite), there is no canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero. Whereas \cite{paper:dirichlet_laplace} used k-means clustering over the posterior samples to perform the actual selection, \cite{bhadra2017horseshoe+} just focused their analysis on the obtained shrinkage and the correct estimation of the sparse signals without any actual selection. In our experiments, we include the DL prior and the regularised horseshoe (RHS) prior \citep{paper:rhs}, which we think scales better to higher number of covariates as compared to other horseshoe-like priors. For the Bayesian regression models with these priors, we compare the posterior shrinkage and the correct estimation of the sparse signals with and without using a reference model as a noise filter. For this purpose, we consider the   average SSE (sum of square errors) and the average SESE (sum of expected square errors) as quantities measuring estimation accuracy. They are defined as
\
\begin{align}
\text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
\text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
&=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}] \label{eq:SESE}
\end{align}
where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$. We divide the errors by $(n-3)$ to have quantities independent of the number of statistical units. Note that the SESE takes the posterior uncertainty in account and thus also the shrinkage of the posterior distributions, whereas the SSE only measures the bias in the point estimates. Note that actually any kind of point estimate could be used to compute the SSE but here we follow \cite{paper:dirichlet_laplace} and use the posterior median. The results that we show later on in Section \ref{shrinkage-signal} are not sensible to such a choice.

For shrinkage priors, one centeral aspect is the amount of shrinkage they actually impose on the relevant and irrelevant features. Ideally, the irrelevant features should be completely shrunken to zero while the relevant features should not be shrunken at all. The amount of shrinkage can be quantitified by the shrinkage factor $\kappa_{j}$ for each parameter $\theta_{j}$ ranging from $0$ (no shrinkage) to $1$ (complete shrinkage). For model \eqref{eq:normal_means_problem2} with a hierarchical shrinkage prior of the type
\begin{align}
Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
&\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
\end{align}
the shrinkage factor $\kappa_{j}$ is defined as
\begin{equation}
\kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}.
\end{equation}
\
In our case, $\kappa_{j}$ can be computed as
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}
\end{equation}
for the regularised horseshoe and the Dirichlet-Laplace prior respectively. In the left hand-side expression, $\tau$ and $\lambda_{j}$ represent the global and local shrinkage parameters of the regularised horseshoe prior \cite[for further details see][]{paper:rhs}, whereas on the right hand-side, $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of the Dirichlet-Laplace prior of parameter $\theta_{j}$ following the notation of \cite{paper:dirichlet_laplace}. We will use the posterior mean of $\kappa_{j}$ as a point estimate.

@@

Here we use simulated data in order to control the difficulty of the selection problem. We rely on the scheme \eqref{eq:simulated_data} using different levels of correlation $\rho\in\{0.3,0.5\}$ and number observation $n\in\{50,70,100\}$. The total number of covariates $p$ and the number of relevant covariates $k$ are fixed to $p = 1000$ and $k = 100$, respectively. In general, the lower $\rho$ and $n$, the more challenging the selection. As mentioned, this data generation mechanism was originally proposed by \cite{paper:projpred} who also devise a reference model performing sufficiently well in this situation. It consists of a linear regression using the first five supervised principal components (SPCs) \citep{bair2006prediction, piironen2018} as covariates and imposing an
hierarchical prior on their coefficients:
\
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10) \\
\end{aligned}
\end{equation}
In the above, $u_{ij}$ represents the $j$-th SPC evaluated at observation $i$, and $s_{max}$ denotes the sample standard deviation of the largest SPC. As screening threshold for the SPCs, we use the value $0.6s_{max}$, which has shown satisfying predictive performance (TODO: clarify this sentence). In our experiments the results are not very sensible to this choice, yet a more principled approach would be to use cross-validation to select the threshold as done in \cite{paper:projpred}.

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of fully Bayesian estimation methods, we analyse the effect of the reference model using two continuous hierarchical shrinkage priors: the Dirichlet-Laplace \citep{paper:dirichlet_laplace} and the regularised horseshoe prior \citep{paper:rhs}. 
% Although the horseshoe+ prior by \cite{bhadra2017horseshoe+} is a state-of-the-art prior for sparse signal analysis in the context of the normal means problem, we decided to include the regularised horseshoe due to its faster MCMC inference, since our purpose in this paper is not to study the best prior choice, but rather compare and show the benefits of using a reference model when carrying out variable selection.
For the Dirichlet hyperparameter of the Dirichlet-Laplace prior, we use an uniform prior on the interval $(1/p,1/2)$. For the reguelarised horseshoe prior, we follow the recommendations provided in \cite{paper:rhs} by setting the prior guess of the number of relevant covariates to one in order to achieve the highest possible shrinkage. We have not observed any sensibility of the results due to such a choice.
\
Since, as explained at the start of Section \ref{comparison}, continuous priors usually do not provide an automatic selection procedure, we study the amount of shrinkage and the bias of the posterior medians looking at the SESE and the SSE (see definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the posterior mean of the shrinkarge factors.

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local false discovery rate \citep{paper:efron, efron2012large}, the empirical Bayes median \citep{johnstone2004needles} and the selection by posterior credible intervals. 
\
The control of the local false discovery rate is applied through the \texttt{R}-package \texttt{locfdr}. It consists of testing the z-values $\{z_{j}\}_{j=1}^{p}$ on whether they belong to the theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$ meaning no relevance) against the alternative hypothesis distribution $f_{1}$. In our case $f_{0}$ corresponds to the standard normal distribution (see expression \eqref{eq:normal_means_problem2}). The quantity of interest is the local false discovery rate (loc.fdr) defined as:
\
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)}
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and $f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal distribution of the z-values. The latter is estimated using splines with 7 degrees of freedom. We select features with local false discovery rate below $0.2$. Such a value is commonly suggested only because it corresponds to a Bayes factor larger than 36 (assuming $\pi_{0}\geq0.9$). Other values can be considered and, in our experience, do not affect the results of the comparison. We use the default setting to estimate $\pi_{0}$ from the data provided by the package.
The empirical Bayes median procedure is given by the \texttt{R}-package \texttt{EbayesThresh} and consists of fitting a Bayesian model with a prior composed by a mixture of a delta in zero and a heavy-tailed distribution. As suggested by (cite), we use a Laplace distribution resulting in a thresholding property, that is, there exists a threshold value such that all the data under that threshold have posterior median equal to zero. Therefore the selection is done by selecting only those parameters whose posterior median is different from zero. The hyperparameter of the Laplace distribution and the mixing weight of the prior are estimated by marginal maximum likelihood. 
The selection by 90\% posterior credible intervals is done using the regularised horseshoe prior and selecting those features whose posterior distribution does not include zero in the interval between the 5\% and the 95\% quantiles.

@@

Figure \ref{fig:stability} shows the estimates of the stability measure proposed by \cite{paper:stability} with 0.95 confidence intervals after 100 simulations. Such a measure takes in account the variability of the subset of the selected features at each simulation (originally at each bootstrap sample), modelling the selection of each variable as a Bernoulli process. Further details are available in (cite). The reference model helps in improving the stability of the selection: Again, such benefit is more notable when the problem is more difficult (small $n$ and $\rho$). In addition, we observe less uncertainty in the stability estimates for the reference approach (i.e., smaller width of the 0.95 confidence intervals), which can be still connected to the overall stability of the procedure.

>>>>

In this section, we demonstrate the benefits of reference model approaches using the normal means problem as a benchmark. The normal means problem consists of estimating the (usually sparse) vector of means of a vector of normally distributed observations. Let us denote the dimensionality of the vector of means by $p$ and let $\{z_{j}\}_{j=1}^{p}$ be the vector of observations of the random variables $\{Z_{j}\}_{j=1}^{p}$. In the normal means problem, the task is to estimate the latent variables $\{\theta_{j}\}_{j=1}^{p}$ of the following model:
\
\begin{equation}\label{eq:normal_means_problem}
Z_{j}|\theta_{j},\sigma\overset{ind}{\sim}N(\theta_{j},\sigma^{2}) \quad j=1,..,p
\end{equation}
Note that it is equivalent to a linear regression where the design matrix is the identity matrix with the number of features being equal to the number of observations. This formulation can be found in practice at various places, 
for example, in the analysis of microarray data, where a large set of genes are tested in two groups of patients labeled as positive or negative to some disease \citep{paper:efron,efron2012large}. The objective of the analysis is to select the subset of genes statistically relevant to the disease. One common way to proceed is to compute the two-samples $t$-statistic for every gene separately.
After normalizing these statistics, they then become the data, $Z$, in the normal means problem \eqref{eq:normal_means_problem}. For further details see the examples in \cite{paper:efron, efron2012large}.

@@

We begin by explaining some Bayesian estimation methods to solving the normal means problem. They all have in common that a Bayesian regression model is fitted using some sparsifying prior on the regression coefficients, for example, as discussed in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. If these priors have continuous support, as is the case for the Dirichlet-Laplace \cite[DL;][]{paper:dirichlet_laplace} or horseshoe-like priors \citep{paper:hs,bhadra2017horseshoe+,paper:rhs}, there is no canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero. Whereas \cite{paper:dirichlet_laplace} used k-means clustering over the posterior samples to perform the actual selection, \cite{bhadra2017horseshoe+} just focused their analysis on the obtained shrinkage and the correct estimation of the sparse signals without any actual selection. In our experiments, we include the DL prior and the regularised horseshoe (RHS) prior \citep{paper:rhs}, which we think scales better to higher number of covariates as compared to other horseshoe-like priors. For the Bayesian regression models with these priors, we compare the posterior shrinkage and the correct estimation of the sparse signals with and without using a reference model as a noise filter. For this purpose, we consider the   average SSE (sum of square errors) and the average SESE (sum of expected square errors) as quantities measuring estimation accuracy. They are defined as
\
\begin{align}
\text{SSE}&=\frac{1}{n-3}\sum_{j=1}^{p}(\hat{\theta}_{j} - \theta^{0}_{j})^{2} \label{eq:SSE} \\
\text{SESE}&=\frac{1}{n-3}\sum_{j=1}^{p}\int_{\mathbb{R}}(\theta_{j}-\theta^{0}_{j})^{2}p(\theta_{j}|z_{1},..,z_{p})d\theta_{j} \\
&=\frac{1}{n-3}\sum_{j=1}^{p}\mathbb{E}_{\theta_{j}}[(\theta_{j}-\theta^{0}_{j})^{2}|z_{1},..,z_{p}] \label{eq:SESE}
\end{align}
where $\theta_{j}^{0}=\sqrt{n-3}\,T_{F}(\rho_{j})$ is the true value, $\hat{\theta}_{j}$ is the posterior median of each parameter $\theta_{j}$. We divide the errors by $(n-3)$ to have quantities independent of the number of statistical units. Note that the SESE takes the posterior uncertainty in account and thus also the shrinkage of the posterior distributions, whereas the SSE only measures the bias in the point estimates. Note that actually any kind of point estimate could be used to compute the SSE but here we follow \cite{paper:dirichlet_laplace} and use the posterior median. The results that we show later on in Section \ref{shrinkage-signal} are not sensible to such a choice.

For shrinkage priors, one central aspect is the amount of shrinkage they actually impose on the relevant and irrelevant features. Ideally, the irrelevant features should be completely shrunken to zero while the relevant features should not be shrunken at all. The amount of shrinkage can be quantified by the shrinkage factor $\kappa_{j}$ for each parameter $\theta_{j}$ ranging from $0$ (no shrinkage) to $1$ (complete shrinkage). For model \eqref{eq:normal_means_problem2} with a hierarchical shrinkage prior of the type
\begin{align}
Z_{j}|&\theta_{j}\overset{ind}{\sim}N(\theta_{j},1) \qquad j=1,..,p \\
&\theta_{j}|\tau,\nu_{j}\overset{ind}{\sim}N(0,\nu_{j}^{2}\tau^{2})
\end{align}
the shrinkage factor $\kappa_{j}$ is defined as
\begin{equation}
\kappa_{j}=\frac{1}{1+\tau^{2}\nu_{j}^{2}}.
\end{equation}
\
In our case, $\kappa_{j}$ can be computed as
\begin{equation}
\kappa_{j}^{\text{RHS}}=\frac{1}{1+\tau^{2}\lambda_{j}^{2}}, \qquad \kappa_{j}^{\text{DL}}=\frac{1}{1+\psi\phi_{j}^{2}\tau^{2}}
\end{equation}
for the regularised horseshoe and the Dirichlet-Laplace prior respectively. In the left hand-side expression, $\tau$ and $\lambda_{j}$ represent the global and local shrinkage parameters of the regularised horseshoe prior \cite[for further details see][]{paper:rhs}, whereas on the right hand-side, $\psi\phi_{j}^{2}\tau^{2}$ is the variance term of the Dirichlet-Laplace prior of parameter $\theta_{j}$ following the notation of \cite{paper:dirichlet_laplace}. We will use the posterior mean of $\kappa_{j}$ as a point estimate.

@@

Here we use simulated data in order to control the difficulty of the selection problem. We rely on the scheme \eqref{eq:simulated_data} using different levels of correlation $\rho\in\{0.3,0.5\}$ and number observation $n\in\{50,70,100\}$. The total number of covariates $p$ and the number of relevant covariates $k$ are fixed to $p = 1000$ and $k = 100$, respectively. In general, the lower $\rho$ and $n$, the more challenging the selection. As mentioned, this data generation mechanism was originally proposed by \cite{paper:projpred} who also devise a reference model performing sufficiently well in this situation. It consists of a linear regression using the first five supervised principal components (SPCs) \citep{bair2006prediction, piironen2018} as covariates and imposing an
hierarchical prior on their coefficients:
\
\begin{equation}
\label{eq:ref_mod}
\begin{aligned}
    Y_{i}|&\boldsymbol{\beta},\sigma^{2},\boldsymbol{u}_{i} \overset{ind}{\sim} N(\boldsymbol{u}_{i}^{T}\boldsymbol{\beta},\sigma^{2}) \quad &i=1,..,n \\
    &\beta_{j}|\!\begin{aligned}[t] &\tau \overset{iid}{\sim} N(0,\tau^{2})\\
    &\tau \sim t_{4}^{+}(0,s_{max}^{-2}) 
    \end{aligned} &j=1,..,5 \\ 
    &\sigma \sim t_{3}^{+}(0,10) \\
\end{aligned}
\end{equation}
In the above, $u_{ij}$ represents the $j$-th SPC evaluated at observation $i$, and $s_{max}$ denotes the sample standard deviation of the largest SPC.
The SPCs are computed through the \texttt{R}-package \texttt{dimreduce} setting the screening threshold parameter at $0.6s_{max}$. 
In our experiments the results are not very sensible to such a choice of the screening threshold, yet a more principled approach would be to use cross-validation to select the threshold as done in \cite{paper:projpred}.

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of fully Bayesian estimation methods, we analyse the effect of the reference model using two continuous hierarchical shrinkage priors: the Dirichlet-Laplace \citep{paper:dirichlet_laplace} and the regularised horseshoe prior \citep{paper:rhs}. 
% Although the horseshoe+ prior by \cite{bhadra2017horseshoe+} is a state-of-the-art prior for sparse signal analysis in the context of the normal means problem, we decided to include the regularised horseshoe due to its faster MCMC inference, since our purpose in this paper is not to study the best prior choice, but rather compare and show the benefits of using a reference model when carrying out variable selection.
For the Dirichlet hyperparameter of the Dirichlet-Laplace prior, we use an uniform prior on the interval $(1/p,1/2)$. For the regularised horseshoe prior, we follow the recommendations provided in \cite{paper:rhs} by setting the prior guess of the number of relevant covariates to one in order to achieve the highest possible shrinkage. We have not observed any sensibility of the results due to such a choice.
\
Since, as explained at the start of Section \ref{comparison}, continuous priors usually do not provide an automatic selection procedure, we study the amount of shrinkage and the bias of the posterior medians looking at the SESE and the SSE (see definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the posterior mean of the shrinkage factors.

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local false discovery rate \citep{paper:efron, efron2012large}, the empirical Bayes median \citep{johnstone2004needles} and the selection by posterior credible intervals. 
\
The control of the local false discovery rate is applied through the \texttt{R}-package \texttt{locfdr}. It consists of testing the z-values $\{z_{j}\}_{j=1}^{p}$ on whether they belong to the theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$ meaning no relevance) against the alternative hypothesis distribution $f_{1}$. In our case $f_{0}$ corresponds to the standard normal distribution (see expression \eqref{eq:normal_means_problem2}). The quantity of interest is the local false discovery rate (loc.fdr) defined as:
\
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)}
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and $f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal distribution of the z-values. The latter is estimated using splines with 7 degrees of freedom. We select features with local false discovery rate below $0.2$. Such a value is commonly suggested only because it corresponds to a Bayes factor larger than 36 (assuming $\pi_{0}\geq0.9$). Other values can be considered and, in our experience, do not affect the results of the comparison. We use the default setting to estimate $\pi_{0}$ from the data provided by the package.
The empirical Bayes median procedure is given by the \texttt{R}-package \texttt{EbayesThresh} and consists of fitting a Bayesian model with a prior composed by a mixture of a delta in zero and a heavy-tailed distribution. As suggested by \cite{johnstone2004needles}, we use a Laplace distribution resulting in a thresholding property, that is, there exists a threshold value such that all the data under that threshold have posterior median equal to zero. Therefore the selection is done by selecting only those parameters whose posterior median is different from zero. The hyperparameter of the Laplace distribution and the mixing weight of the prior are estimated by marginal maximum likelihood. 
The selection by 90\% posterior credible intervals is done using the regularised horseshoe prior and selecting those features whose posterior distribution does not include zero in the interval between the 5\% and the 95\% quantiles.

@@

Figure \ref{fig:stability} shows the estimates of the stability measure proposed by \cite{paper:stability} with 0.95 confidence intervals after 100 simulations. Such a measure takes in account the variability of the subset of the selected features at each simulation (originally at each bootstrap sample), modelling the selection of each variable as a Bernoulli process. Further details are available in \cite{paper:stability}. The reference model helps in improving the stability of the selection: Again, such benefit is more notable when the problem is more difficult (small $n$ and $\rho$). In addition, we observe less uncertainty in the stability estimates for the reference approach (i.e., smaller width of the 0.95 confidence intervals), which can be still connected to the overall stability of the procedure.
